---
title: "Analysis Of Influential Valriables"
author: "Himangshu Pal"
date: "February 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Libraries**


```{r}
library(lubridate)

library(vcd)

library(ggplot2)

library(pastecs)

library(gmodels)

library(boot)

library(corrplot)

library(car)

library(tree)

library(pls)

library(randomForest)



```

__Read input file.__ 
```{r}

myData <- read.csv("house_country.csv", header = T)

```

__Explore Data__


```{r}

str(myData)

head(myData)

names(myData)

```

__Let's work with the date field.__

```{r}
myDate <- myData$date

myDate <- as.Date(myDate, "%m/%d/%Y")

myMonth <- month(myDate, label = T)

myMonth <- factor(myMonth, ordered = FALSE) 


myMonth <- factor(myMonth)

myYear <- year(myDate)

myYear <- factor(myYear)

```



```{r}
# housePrice is the new data frame with year and month, minus housing price index and date

housePrice <- data.frame(obs_year = myYear,       
                         obs_month = myMonth,
                         myData[,-c(1,19)])


sum(is.na(housePrice))   # No missing value


```


__Time Series Analysis__

```{r}

# Let's check how each parameter behave in time. 

start_yr <- min(year(myDate))

end_yr <- max(year(myDate)) 

ts_obj1<- ts(housePrice$median_house_price, start = c(start_yr, 1), frequency = 4)

plot(ts_obj1,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Median House Price", main = "Year vs Price")


ts_obj2<- ts(housePrice$NASDAQ_comp_index, start = c(start_yr, 1), frequency = 4)

plot(ts_obj2,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Stock Index", main = "Year vs Stock Market")


ts_obj3<- ts(housePrice$total_unemployed, start = c(start_yr, 1), frequency = 4)

plot(ts_obj3,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Unemployment", main = "Year vs Unemployement")


ts_obj4<- ts(housePrice$job_loser_rate, start = c(start_yr, 1), frequency = 4)

plot(ts_obj4,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Job Loser", main = "Year vs Job Loser")


ts_obj5<- ts(housePrice$multi_job_holders, start = c(start_yr, 1), frequency = 4)

plot(ts_obj5,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Multiple Job Holder", main = "Year vs Multi Job Holder")


ts_obj6<- ts(housePrice$not_in_labor_searched_for_work, start = c(start_yr, 1), frequency = 4)

plot(ts_obj6,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Not in labor pool", main = "Year vs Not in labor pool")


ts_obj7<- ts(housePrice$job_leaver, start = c(start_yr, 1), frequency = 4)

plot(ts_obj7,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Job Leavers", main = "Year vs Job Leavers")


ts_obj8<- ts(housePrice$more_than_15_weeks, start = c(start_yr, 1), frequency = 4)

plot(ts_obj8,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "No Job > 15 weeks", main = "Year vs Jobless > 15 weeks")


ts_obj9<- ts(housePrice$disp_income, start = c(start_yr, 1), frequency = 4)

plot(ts_obj9,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Disposable", main = "Year vs Disposable Income")


ts_obj10<- ts(housePrice$FRB_Rates, start = c(start_yr, 1), frequency = 4)

plot(ts_obj10,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "FRB Rates", main = "Year vs FRB Rates")


ts_obj11<- ts(housePrice$gross_dom_prod, start = c(start_yr, 1), frequency = 4)

plot(ts_obj11,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "GDP", main = "Year vs GDP")


ts_obj12<- ts(housePrice$consumer_price_index, start = c(start_yr, 1), frequency = 4)

plot(ts_obj12,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "CPI", main = "Year vs CPI")


ts_obj13<- ts(housePrice$prod_price_index, start = c(start_yr, 1), frequency = 4)

plot(ts_obj13,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "PPI", main = "Year vs PPI")


ts_obj14<- ts(housePrice$long_interest_rate, start = c(start_yr, 1), frequency = 4)

plot(ts_obj14,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Long Interest Rate", main = "Year vs Long Interest Rate")


ts_obj15<- ts(housePrice$mortgage_Rate30, start = c(start_yr, 1), frequency = 4)

plot(ts_obj15,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "30 years Mortgage", main = "Year vs Mortgage Rate 30 yr")


ts_obj16<- ts(housePrice$mortgage_Rate15, start = c(start_yr, 1), frequency = 4)

plot(ts_obj16,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "15 years Mortgage", main = "Year vs Mortgage Rate 15 yr")


ts_obj17<- ts(housePrice$population, start = c(start_yr, 1), frequency = 4)

plot(ts_obj17,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Population", main = "Year vs Population")


ts_obj18<- ts(housePrice$total_expenditures, start = c(start_yr, 1), frequency = 4)

plot(ts_obj18,type= "l", lwd = 2, col = "blue", xlab = "Year", ylab = "Individual Expenditures", main = "Year vs Individual Expenditures")

```


__Let's bring up some plots and graphs to find some pattern.__

```{r}

# Response is median_house_price. So, she is the center attraction of all plots. 

hist(housePrice$median_house_price, 
     breaks = 15,
     col = "grey",
     xlab = "Median House Price",
     main = "Histogram of Median House Price")
box()


hist(housePrice$median_house_price, 
     freq = FALSE,
     breaks = 15,
     col = "grey",
     xlab = "Median House Price",
     main = "Histogram, rug plot, density curve")
rug(jitter(housePrice$median_house_price))
lines(density(housePrice$median_house_price), col = "red", lwd = 1)
box()


boxplot(housePrice$median_house_price ~ factor(housePrice$obs_month, ordered = F), 
        varwidth = T, col = "grey", main = "Box Plot Median House Price Per Quater", xlab = "Month",
        ylab = "Median House Price")
box()
```



__Response variable's Relationship with Predictors and also verify the linearity__

```{r}

ggplot(data = housePrice, 
       aes(x = population, y = median_house_price)) +
    geom_point(pch=20, size = 2) +
    geom_smooth(method = "lm", formula = y ~ x, color = "blue", linetype = 14) +
    labs(title = "Population vs median house price", 
         x = "Population", y = "Median House Price")
# Relationship is non-linear between Population and Median house price 

ggplot(data = housePrice,
       aes(x = NASDAQ_comp_index, y = median_house_price)) +
    geom_smooth(method = "lm", formula = y ~ x, color = "blue", linetype = 2) + #, formula = y ~ I(x ^ -2),) +
    geom_point(pch = 15, size = 1) +
    labs(title = "Stock Index vs median house price", 
         x = "NASDAQ-Index", y = "Median House Price")



```



```{r}

ggplot(data = housePrice,
       aes(x = consumer_price_index, y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x, color = "blue", linetype = 2) +  #   ,formula = y ~ I(x ^ -2),
    geom_point(size = 2) +
    labs(title = "CPI vs median house price", 
         x = "CPI", y = "Median House Price")

    

ggplot(data = housePrice,
       aes(x = prod_price_index, y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x , color = "blue", linetype = 2) +
    geom_point(size = 2) +
    labs(title = "PPI vs median house price", 
         x = "PPI", y = "Median House Price")

    


ggplot(data = housePrice,
       aes(x = gross_dom_prod , y = median_house_price)) +
    geom_smooth(method = lm,  formula = y ~ x, color = "blue", linetype = 2) + 
    geom_point(size = 2) +
    labs(title = "GDP vs median house price", x = "GDP", y = "Median House Price")


```


```{r}

ggplot(data = housePrice,
       aes(x = FRB_Rates, y = median_house_price)) +
    geom_smooth(method = lm, formula = (y ~ x), color = "blue", linetype = 2) + #formula = y ~ poly(x , 9),
    geom_point(size = 2) +
    labs(title = "FRB_Rates vs median house price", 
         x = "FRB_Rates", y = "Median House Price")



ggplot(data = housePrice,
       aes(x = long_interest_rate  , y = median_house_price)) +
    geom_smooth(method = lm, formula = (y ~ x), color = "blue", linetype = 2) +
    geom_point(size = 2)    +
    labs(title = "Long Interest Rates vs median house price", 
         x = "Interest Rates", y = "Median House Price")



ggplot(data = housePrice,
       aes(x = mortgage_Rate30, y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x, color = "blue", linetype = 2) +
    geom_point(size = 2)    +
    labs(title = "Mortgage rate 30 yr vs median house price", 
         x = "Mortgare rate 30 yr", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = mortgage_Rate15 , y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x, color = "blue", linetype = 2) +
    geom_point(size = 2)    +
    labs(title = "Mortgage rate 15 yr vs median house price", 
         x = "Mortgage rate 15 yr", y = "Median House Price")

```



```{r}
ggplot(data = housePrice,
       aes(x = total_expenditures, y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x, color = "blue", linetype = 2) +
    geom_point(size = 2) +
    labs(title = "total expenditures vs median house price", 
         x = "total_expenditures", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = disp_income, y = median_house_price)) +
    geom_smooth(method = lm, formula = y ~ x, color = "blue", linetype = 2) +
    geom_point(size = 2)    +
    labs(title = "Disposable Income vs median house price", 
         x = "Disposable Income", y = "Median House Price")

```

__Employement__

```{r}

ggplot(data = housePrice,
       aes(x = multi_job_holders, y = median_house_price)) +
       geom_smooth(method = lm, formula = y ~ x, se = F, color = "blue", linetype = 3) +
       geom_point(size = 2) +
       labs(title = "Multi job holders vs median house price", 
         x = "Multi Job Holders", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = job_loser_rate , y = median_house_price)) +
       geom_smooth(method = lm, formula = y ~ x , se = T, color = "blue", linetype = 3) +
       geom_point(size = 2) +
       labs(title = "Job Loser Rate vs median house price", 
         x = "Job Loser Rate", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = more_than_15_weeks, y = median_house_price)) +
       geom_smooth(method = lm, formula = y ~ x, se = F, color = "blue", linetype = 3) +
       geom_point(size = 2) +
       labs(title = "Job less > 15 weeks vs median house price", 
         x = "Job less > 15 weeks", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = not_in_labor_searched_for_work, y = median_house_price)) +
      geom_smooth(method = lm, formula = y ~ x, se = FALSE, size = 1, color = "blue", linetype = 3) +
      geom_point(size = 2)    +
      labs(title = "Not in labor pool vs median house price", 
         x = "Not in labor pool ", y = "Median House Price")


ggplot(data = housePrice,
       aes(x = total_unemployed , y = median_house_price)) +
       geom_smooth(method = lm, formula = y ~  x, se = F, color = "blue", linetype = 3) +
       geom_point(size = 2) +
       labs(title = "Unemployed vs median house price", 
         x = "Unemployed ", y = "Median House Price")



ggplot(data = housePrice,
       aes(x = job_leaver, y = median_house_price)) +
       geom_smooth(method = lm,formula = y ~ x , se = FALSE, size = 1, color = "blue", linetype = 3) +
       geom_point(size = 2)    +
       labs(title = "  Job leaver vs median house price", 
         x = " Job leaver ", y = "Median House Price")



```


__Descriptive Analysis__

```{r}
# Let's look at measures of central tendency, variablility and destribution shape of continuous variables. 

# Summary 

summary(housePrice)

```

__Standardize data__

```{r}

myData1 = housePrice[,-c(1:2)]  # No need of year and month

#myData1 = housePrice

myData1= scale(myData1, center =T, scale = T)

housePrice_scale = myData1

housePrice1 = myData1

housePrice1 = as.data.frame(housePrice1)

names(housePrice1) <- c("cpi", "ppi", "income", "gdp", "frb","expence","interest", "mrate15","mrate30","stock","unemp","looser","multijob", "nojobsearch","leaver","nojob15","popu","medHouse")

head(housePrice1)

```

__Correlation Matrix__

```{r}
m <- cor(housePrice_scale)

rownames(m) <- c("cpi", "ppi", "income", "gdp", "frb","expence","interest", "mrate15","mrate30","stock","unemp","looser","multijob", "nojobsearch","leaver","nojob15","popu","medHouse")

colnames(m) <- c("cpi", "ppi", "income", "gdp", "frb","expence","interest", "mrate15","mrate30","stock","unemp","looser","multijob", "nojobsearch","leaver","nojob15","popu","medHouse")



corrplot.mixed(m, lower.col = "black", number.cex = .7, tl.cex=.7 )

```



__Random Sample for Training and Test sets__

```{r}

head(housePrice1)

attach(housePrice1)

#x = model.matrix(medHouse ~., housePrice1)[,-1]
#y = housePrice1$medHouse


set.seed(1)

#train = sample(1:nrow(x), round(nrow(x) * .75))

train_index = sample(1:nrow(housePrice1), round(nrow(housePrice1) * .75))

train = housePrice1[train_index,]
test = housePrice1[-train_index,-18]
#y.test = y[test]
y.test = housePrice1[-train_index, 18]

```

__Subset Selection For Linear Regression__

```{r}
# PCR

## I need to FIX PCR

#set.seed(1)

#pcr.fit = pcr(medHouse ~ . , data = housePrice1, subset = train, scale = TRUE, validation = "CV")

set.seed(1)

pcr.fit = pcr(medHouse ~ . , data = housePrice1,  scale = TRUE, validation = "CV")

validationplot(pcr.fit, val.type = "R2")

summary(pcr.fit)

#dimnames(housePrice1)
#pcr.pred = predict(pcr.fit, x[test,], ncomp = 10)

pcr.fit = pcr(medHouse~., data = train, scale = TRUE, validation = "CV")

pcr.pred = predict(pcr.fit, data = test, ncomp = 2)

mean((pcr.pred - y.test) ^ 2)



#pcr.fit = pcr(y ~ x, scale = TRUE, ncomp = 10)

pcr.fit = pcr(y ~ x, data = scale = TRUE, ncomp = 2)

pcr.fit$coefficients


summary(pcr.fit)

mean((pcr.pred - y.test) ^ 2)

par(mfrow = c(3,1))
validationplot(pcr.fit)
validationplot(pcr.fit, val.type = "MSEP")
validationplot(pcr.fit, val.type = "R2")

predplot(pcr.fit)

coefplot(pcr.fit)

```


```{r}
# PLS

set.seed(1)

pls.fit = plsr(medHouse ~. , data = housePrice1, subset = train, scale = TRUE, validation = "CV")

summary(pls.fit)

# M =13, 99.97% 

pls.pred = predict(pls.fit, x[test,], ncomp = 10)

mean((pls.pred - y.test) ^ 2)

#Full data set

pls.fit = plsr(medHouse ~ . , data = housePrice1, scale = TRUE, ncomp =10)

summary(pls.fit)

plot(pls.fit)

par(mfcol = c(3,1))

validationplot(pls.fit)
validationplot(pls.fit, val.type = "MSEP")
validationplot(pls.fit, val.type = "R2")

```

__Non-linear regression__

```{r}

set.seed(1)

tree.house = tree(medHouse ~.,housePrice1, subset = train)

summary(tree.house)

plot(tree.house)

text(tree.house, pretty = 0)


#prunning the tree
cv.house = cv.tree(tree.house)

plot(cv.house$size, cv.house$dev, type = "b")

#minimum error occured for tree size 5.

prune.house = prune.tree(tree.house, best = 5) 

plot(prune.house)

text(prune.house, pretty = 0)


#Prediction and validation 
yhat = predict(tree.house, newdata = housePrice1[-train,])

house.test = housePrice1[-train, "medHouse"]

plot(yhat, house.test)
abline(0,1)

mean((yhat - house.test)^2)

#0.03445708

```

__Bagging__


```{r}

set.seed(1)

bag.house = randomForest(medHouse ~., data = housePrice1, subset = train, mtry = ncol(housePrice1) - 1, importance = T)

importance(bag.house)

#cpi         11.097628   10.38535960
#ppi         13.091498   13.70898383
#gdp         12.867463   11.50522481

#plot(bag.house)
#text(bag.house, pretty = 0)

yhat.bag = predict(bag.house, newdata = housePrice1[-train,])

plot(yhat.bag, house.test)
abline(0,1)

mean((yhat.bag - house.test)^2)


#0.009225094
#cpi         0.2005510032   10.46006450
#ppi         0.2712617087   14.48534045


#Again bagging using ntree option

bag.house = randomForest(medHouse~.,data = housePrice1, subset = train, mtry = ncol(housePrice1)-1, ntree = 25)

yhat.bag = predict(bag.house, newdata = housePrice1[-train,])

mean((yhat.bag - house.test) ^2)

importance(bag.house)
#0.01088559
#0.0124372

```



__Random Forest__

```{r}
set.seed(1)

rf.house = randomForest(medHouse~.,data = housePrice1, subset = train, mtry = 6, importance = T)

plot(rf.house)

yhat.rf = predict(rf.house, newdata = housePrice1[-train,]) 

mean((yhat.rf - house.test) ^ 2)    #MSE

#0.008968475
#0.009251125

importance(rf.house)

#cpi         12.320385   11.90178430
#income      11.257960   10.39671223
#gdp         12.510731   12.01185450
#popu        12.568885   10.92544258

varImpPlot(rf.house, pch = 20)
```


__Boosting__


```{r}
library(gbm)

set.seed(1)

boost.house = gbm(medHouse~., data = housePrice1[train,], distribution = "gaussian",
                  n.tree = 5000, interaction.depth = 4)

summary(boost.house)

par(mfrow = c(1,2))
plot(boost.house, i = "cpi")
plot(boost.house, i = "ppi")

#Prediction 

yhat.boost = predict(boost.house, newdata = housePrice1[-train,], n.tree = 5000)

mean((yhat.boost - house.test) ^ 2)

#0.01115614


#Boosting with shrinkage (lambda) = 0.2

boost.house = gbm(medHouse~., data = housePrice1[train,], distribution = "gaussian",
                     n.trees = 5000, interaction.depth = 4, shrinkage = .2, verbose = F)

yhat.boost = predict(boost.house, newdata = housePrice1[-train, ], n.trees = 5000)

mean((yhat.boost - house.test)^2)

#0.0163058
```

__Dimension Reduction__


```{r}

# PCA 


pr.out = prcomp(housePrice1[,-18], scale = F) 

print(pr.out)

#names(pr.out)

summary(pr.out)

plot(pr.out, type = "l")

# center = mean
# scale = SD

pr.out$center

pr.out$scale

pr.out$rotation

dim(pr.out$x)

biplot(pr.out, scale = 0)

#?biplot

pr.out$rotation = -pr.out$rotation

pr.out$x = -pr.out$x

biplot(pr.out, scale = 0,cex = .9 )

print(pr.out$rotation)
#?biplot

pr.out$sdev

pr.var = pr.out$sdev ^ 2
pr.var


pve = pr.var/sum(pr.var)



pve

par(mfrow = c(1,2))

plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained",
     ylim = c(0,1), type = "b")

plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0,1), type = "b")




```

